HUME: Human UCCA-Based Evaluation of Machine Translation

### 2016-03-18 paper/main.tex
Commented out older graphics and heatmaps and removed comments to see how much space we now have

\newcommand{\com}[1]{}
\newcommand{\secref}[1]{Section~\ref{#1}}
\newcommand{\figref}[1]{Figure~\ref{#1}}
\newcommand{\tabref}[1]{Table~\ref{#1}}
\newcommand{\XXX}[1]{{\color{red}XXX #1}} % a general todo or comment
\newcommand{\oa}[1]{\footnote{\color{red}OA: #1}}
\newcommand{\bh}[1]{\footnote{\color{blue}BH: #1}}

@@

\def\equo#1{``{#1}''}  % curly quotes

@@

  Recent interest in semantics-aware MT systems underscores the importance of
  semantic evaluation for Machine Translation (MT) systems. 
  A semantic measure, sensitive to possible semantic discrepancies
  between the translation output and the source, is likely support the construction
  and tuning of semantics-based MT, much in the same way that string-based measures,
  such as BLEU, enabled progress in shallower MT.
  We present a novel human semantic evaluation measure for MT, Human
  UCCA-based MT Evaluation (HUME), building on the UCCA semantic representation scheme.
  Making progress over previous work, the proposed measure takes into account
  a wider range of semantic phenomena and does not rely on semantic annotation
  of the MT output itself, whose interpretation is often unclear.
  We experiment with four language pairs, demonstrating HUME's broad applicability,
  as well as its reliability, reflected in its inter-annotator agreement rates.

@@

\begin{figure}

\begin{tikzpicture}[semithick, x=2.4cm,y=0.8cm, >=latex, baseline=7ex,
    inner sep=.3ex, outer sep=.7ex, minimum size=1.2ex,->]
  \node (ROOT) [fill=black, circle] {}
  child {node {After} edge from parent node[left]  {$L$\hspace{1.5cm}}}
  child {node (H1)  [fill=black,circle] {}
    child {node {graduation} edge from parent node[right] {$P$}}
      %child {node {to Paris} edge from parent node[right] {$A$}}
      edge from parent node[left] {$H$}
    }
    child {node (H2)  [fill=black,circle] {}
      child {node  {} edge from parent[draw=none]}
      child {node  {} edge from parent[draw=none]}
      child {node  (Tom) {Tom} edge from parent node[right] {$A$}}
      child {node  {moved} edge from parent node[right] {$P$}}
      child {node  [fill=black,circle] {}
        child {node  {to} edge from parent node[right] {$R$}}
        child {node  {NYC} edge from parent node[right] {$C$}}
        edge from parent node[right] {$A$}}
      %child {node {to Paris} edge from parent node[right] {$A$}}
      edge from parent node[left] {$H$}
    };
    \draw (H1) -> (Tom) node[midway, right] {$A$ \hspace{0.2cm}};
  \end{tikzpicture}
\caption{\label{fig:ucca_example}
  Sample UCCA annotation of the sentence ``After graduation, Tom moved
  to NYC''. Leaves correspond to words, whereas nodes correspond to units.
  Edge labels mark UCCA categories, which are described in the table.}
%which are largely ignored in this work
\end{figure}

@@

\begin{figure*}
  \begin{adjustbox}{trim=2.3cm 0 0 0}
   \begin{tikzpicture}
     [ level distance=12mm,
	   level 1/.style={sibling distance=7.5em,level distance=7mm},
       level 2/.style={sibling distance=5em,level distance=12mm},
       level 3/.style={sibling distance=5em},
       level 4/.style={sibling distance=3em}]
     %[semithick, x=2.4cm,y=0.8cm, >=latex, baseline=7ex,
     %  inner sep=.3ex, outer sep=.7ex, minimum size=1.2ex,->]
   \node (ROOT) [label={B}] [fill=blue, circle] {}
   child {node (H3) [label={A}] [fill=yellow,circle] {}
     child {node [label={[red]:R}] {Golf} edge from parent node[left] {}}
     child {node (H1) [label={[orange]30:O}] [fill=orange,circle] {}
       child {node [label={}] {became} edge from parent node[left]  {}}
       child {node [label={}] {a} edge from parent node[left]  {}}
       child {node [label={}] {passion} edge from parent node[left]  {}}
     }
     child {node (H2) [label={A}] [fill=yellow,circle] {}
       child {node {} edge from parent[draw=none] {}}
       child {node {} edge from parent[draw=none] {}}
       child {node {} edge from parent[draw=none] {}}
       child {node {} edge from parent[draw=none] {}}
       child {node [label={[red]:R}] {for} edge from parent node[left]  {}}
       child {node [label={[green]:G}] {his} edge from parent node[left]  {}}
       child {node [label={[orange]:O}] {oldest} edge from parent node[left]  {}}
       child {node [label={[green]:G}] {daughter} edge from parent node[left]  {}}
     }
   }
   child {node {} edge from parent[draw=none] {}}
   child {node (H4) [label={[]30:A}] [fill=yellow,circle] {}
     child {node (she) [label={[green]:G}] {she} edge from parent node[left]  {}}
     child {node [label={[green]:G}] {attended} edge from parent node[left]  {}}
     child {node [label={[green]:G}] {daily} edge from parent node[left]  {}}
     child {node [label={[green]:G}] {lessons} edge from parent node[left]  {}}
   }
   child {node [label={[red]:R}] {and} edge from parent node[left]  {}}
   child {node (H5) [label={A}] [fill=yellow,circle] {}  
     child {node [label={[red]:R}] {greatly} edge from parent node[left] {}}
     child {node [label={[orange]:O}] {improved} edge from parent node[left] {}}
   };
   \draw[dashed] (H5) -> (she) ;

   \com{
   child {node (H1) [label={\#8}] [fill=black,circle] {}
     child {node [label={\#2}] {graduation} edge from parent node[right] {$P$}}
       %child {node {to Paris} edge from parent node[right] {$A$}}
       edge from parent node[left] {$H$}
     }
     child {node (H2) [label={\#9}] [fill=black,circle] {}
       child {node  {} edge from parent[draw=none]}
       child {node  {} edge from parent[draw=none]}
       child {node [label={\#3}] (Tom) {Tom} edge from parent node[right] {$A$}}
       child {node [label={\#4}] {moved} edge from parent node[right] {$P$}}
       child {node [label={\#10}] [fill=black,circle] {}
         child {node [label={\#5}] {to} edge from parent node[right] {$R$}}
         child {node [label={\#6}] {NYC} edge from parent node[right] {$C$}}
         edge from parent node[right] {$A$}}
       %child {node {to Paris} edge from parent node[right] {$A$}}
       edge from parent node[left] {$H$}
     };
     \draw (H1) -> (Tom) node[midway, right] {$A$ \hspace{0.2cm}};
     }
   \end{tikzpicture}
  \\
  \end{adjustbox}
    % Ondrej removed the width=\textwidth from the options below to keep font
	% unenlarged:
  \adjustbox{center,margin=.5em}{\it His old daughter loved tennis, and took daily lessons however good he was}

  \caption{\label{fig:hume_tree}
    Sample UCCA annotation of a source sentence, translation (below in italics), 
    and translation judgements over the source units, marked with A,B,R,O,G labels. Yellow/blue marks
    nodes annotated with A/B, respectively.
    In our experiments, the translation is in German, Romaninan, Polish or Czech.
    ``became a passion'' is marked as atomic, as it is a multi-word expression.
    The dashed line indicates that ``she'' is also a child of the unit ``greatly improved''.
    UCCA categories, and alignment between source and translation are omitted for brevity.}

\end{figure*}

@@

It is also relatively common in translation that verbal constructions get
nominalized or vice versa. An example from our dataset:
  ``Two {\bf review} authors assessed the studies independently\dots{}'' --
  ``Zwei Autoren bewertete die Studien unabh\"angig zu {\bf
  \"uberpr\"ufen}\dots{}'' (lit. ``Two authors assessed the studies
  independently to be reviewed\dots").\XXX{This is Omri's example, with
  Ondrej's gloss for German. Ondrej is
  not totally sure that the meaning is well preserved.}

Here is \XXX{another} example with a missing linker: the source \equo{However,
this review\dots} was translated as \equo{Diese
Überprüfung\dots} (lit. \equo{This review}). The omission is not by HMEANT at
all.

@@

\paragraph{Annotation of Predicates}

HUME allows to annotate multi-word and even
non-contiguous elements as a single unit. HMEANT, on the other hand, always
required exactly one word to be annotated as the predicate (``Action''), which
is often linguistically inadequate.

Another problem with HMEANT identified by \perscite{bojar:wu:ssst:2012} was the
inability to indicate that the predicate is somewhat mistranslated (e.g.
accidentally negated or in a slightly wrong tense). HMEANT
annotation allows either to mark the source and reference frames as matching
(which implies that the corresponding predicates are correct translations of each
other), or as non-matching (which leads to a zero score for the whole frame) but
nothing in between. In HUME, the verb is still present as a sub-unit of the
frame and has the traffic-lights annotation.

\XXX{Omri: I would consider adding further examples from the Google Doc. For instnace
``your risk of falling'' -- ``Ihr Risiko''.
Ondrej: I don't see any HMEANT problem here, the whole NP 'your risk of falling'
is aligned with 'Ihr Risiko' and HMEANT allows to indicate that the match was
partial, to that's probably what everyone would choose. I've put the other
example higher, but please have a look, I'm not sure if we should keep it there
at all.}

@@

}
\end{tabular}
\caption{Confusion matrices for each of the language pairs, plotted as heatmaps.}
\label{fig:heatmap}
\end{figure*}

>>>>

\newcommand{\com}[1]{}
\newcommand{\secref}[1]{Section~\ref{#1}}
\newcommand{\figref}[1]{Figure~\ref{#1}}
\newcommand{\tabref}[1]{Table~\ref{#1}}
\newcommand{\XXX}[1]{{\color{red}XXX #1}} % a general todo or comment
%\newcommand{\oa}[1]{\footnote{\color{red}OA: #1}}
\newcommand{\oa}[1]{}
%\newcommand{\bh}[1]{\footnote{\color{blue}BH: #1}}
\newcommand{\bh}[1]{}

@@

@@

%  Recent interest in semantics-aware MT systems underscores the importance of
%  semantic evaluation for Machine Translation (MT) systems. 
%  A semantic measure, sensitive to possible semantic discrepancies
%  between the translation output and the source, is likely to support the construction
%  and tuning of semantics-based MT, much in the same way that string-based measures,
%  such as BLEU, enabled progress in shallower MT.
%  We present a novel human semantic evaluation measure for MT, Human
%  UCCA-based MT Evaluation (HUME), building on the UCCA semantic representation scheme.
%  Making progress over previous work, the proposed measure takes into account
%  a wider range of semantic phenomena and does not rely on semantic annotation
%  of the MT output itself, whose interpretation is often unclear.
%  We experiment with four language pairs, demonstrating HUME's broad applicability,
%  as well as its reliability, reflected in its inter-annotator agreement rates.

Human evaluation of machine translation is a cognitively difficult task and when annotators rate or rank complete sentences we see low levels of agreement especially
for longer sentences.  
In order to extract more reliable annotations, previous research broke down evaluation into smaller components by asking humans to annotate semantic roles on the source and translation, and then to align them. 
We propose a new evaluation method which is based on semantic annotation called Universal Conceptual Cognitive Annotation(UCCA). UCCA is applicable across many languages, it provides a complete structure over sentences, and it is grounded in the words in the sentence. 
We apply UCCA only over the source or reference sentence, and not over the possibly garbled translation, 
and this makes annotation more reliable and also efficient as we can reuse labor intensive semantic annotations. We show that HUME can be reliably applied to four different target languages and that sentence level inter annotor agreement does not decrease over longer sentences.

@@

%\begin{figure}
%
%\begin{tikzpicture}[semithick, x=2.4cm,y=0.8cm, >=latex, baseline=7ex,
%    inner sep=.3ex, outer sep=.7ex, minimum size=1.2ex,->]
%  \node (ROOT) [fill=black, circle] {}
%  child {node {After} edge from parent node[left]  {$L$\hspace{1.5cm}}}
%  child {node (H1)  [fill=black,circle] {}
%    child {node {graduation} edge from parent node[right] {$P$}}
%      %child {node {to Paris} edge from parent node[right] {$A$}}
%      edge from parent node[left] {$H$}
%    }
%    child {node (H2)  [fill=black,circle] {}
%      child {node  {} edge from parent[draw=none]}
%      child {node  {} edge from parent[draw=none]}
%      child {node  (Tom) {Tom} edge from parent node[right] {$A$}}
%      child {node  {moved} edge from parent node[right] {$P$}}
%      child {node  [fill=black,circle] {}
%        child {node  {to} edge from parent node[right] {$R$}}
%        child {node  {NYC} edge from parent node[right] {$C$}}
%        edge from parent node[right] {$A$}}
%      %child {node {to Paris} edge from parent node[right] {$A$}}
%      edge from parent node[left] {$H$}
%    };
%    \draw (H1) -> (Tom) node[midway, right] {$A$ \hspace{0.2cm}};
%  \end{tikzpicture}
%\caption{\label{fig:ucca_example}
%  Sample UCCA annotation of the sentence ``After graduation, Tom moved
%  to NYC''. Leaves correspond to words, whereas nodes correspond to units.
%  Edge labels mark UCCA categories, which are described in the table.}
%%which are largely ignored in this work
%\end{figure}

@@

%\begin{figure*}
%  \begin{adjustbox}{trim=2.3cm 0 0 0}
%   \begin{tikzpicture}
%     [level 1/.style={sibling distance=7.5em},
%       level 2/.style={sibling distance=5em},
%       level 3/.style={sibling distance=5em},
%       level 4/.style={sibling distance=3em}]
%     %[semithick, x=2.4cm,y=0.8cm, >=latex, baseline=7ex,
%     %  inner sep=.3ex, outer sep=.7ex, minimum size=1.2ex,->]
%   \node (ROOT) [label={B}] [fill=blue, circle] {}
%   child {node (H3) [label={A}] [fill=yellow,circle] {}
%     child {node [label={[red]:R}] {Golf} edge from parent node[left] {}}
%     child {node (H1) [label={[orange]30:O}] [fill=orange,circle] {}
%       child {node [label={}] {became} edge from parent node[left]  {}}
%       child {node [label={}] {a} edge from parent node[left]  {}}
%       child {node [label={}] {passion} edge from parent node[left]  {}}
%     }
%     child {node (H2) [label={A}] [fill=yellow,circle] {}
%       child {node {} edge from parent[draw=none] {}}
%       child {node {} edge from parent[draw=none] {}}
%       child {node {} edge from parent[draw=none] {}}
%       child {node {} edge from parent[draw=none] {}}
%       child {node [label={[red]:R}] {for} edge from parent node[left]  {}}
%       child {node [label={[green]:G}] {his} edge from parent node[left]  {}}
%       child {node [label={[orange]:O}] {oldest} edge from parent node[left]  {}}
%       child {node [label={[green]:G}] {daughter} edge from parent node[left]  {}}
%     }
%   }
%   child {node {} edge from parent[draw=none] {}}
%   child {node (H4) [label={[]30:A}] [fill=yellow,circle] {}
%     child {node (she) [label={[green]:G}] {she} edge from parent node[left]  {}}
%     child {node [label={[green]:G}] {attended} edge from parent node[left]  {}}
%     child {node [label={[green]:G}] {daily} edge from parent node[left]  {}}
%     child {node [label={[green]:G}] {lessons} edge from parent node[left]  {}}
%   }
%   child {node [label={[red]:R}] {and} edge from parent node[left]  {}}
%   child {node (H5) [label={A}] [fill=yellow,circle] {}  
%     child {node [label={[red]:R}] {greatly} edge from parent node[left] {}}
%     child {node [label={[orange]:O}] {improved} edge from parent node[left] {}}
%   };
%   \draw[dashed] (H5) -> (she) ;
% 
%   \com{
%   child {node (H1) [label={\#8}] [fill=black,circle] {}
%     child {node [label={\#2}] {graduation} edge from parent node[right] {$P$}}
%       %child {node {to Paris} edge from parent node[right] {$A$}}
%       edge from parent node[left] {$H$}
%     }
%     child {node (H2) [label={\#9}] [fill=black,circle] {}
%       child {node  {} edge from parent[draw=none]}
%       child {node  {} edge from parent[draw=none]}
%       child {node [label={\#3}] (Tom) {Tom} edge from parent node[right] {$A$}}
%       child {node [label={\#4}] {moved} edge from parent node[right] {$P$}}
%       child {node [label={\#10}] [fill=black,circle] {}
%         child {node [label={\#5}] {to} edge from parent node[right] {$R$}}
%         child {node [label={\#6}] {NYC} edge from parent node[right] {$C$}}
%         edge from parent node[right] {$A$}}
%       %child {node {to Paris} edge from parent node[right] {$A$}}
%       edge from parent node[left] {$H$}
%     };
%     \draw (H1) -> (Tom) node[midway, right] {$A$ \hspace{0.2cm}};
%     }
%   \end{tikzpicture}
%  \\
%  \end{adjustbox}
%%  \adjustbox{width=\textwidth,center,margin=.5em}{\it His old daughter loved tennis, and took daily lessons however good he was}
%
%  \caption{\label{fig:hume_tree}
%    Sample UCCA annotation of a source sentence, translation (below in italics), 
%    and translation judgements over the source units, marked with A,B,R,O,G labels. Yellow/blue marks
%    nodes annotated with A/B, respectively.
%    In our experiments, the translation is in German, Romaninan, Polish or Czech.
%    ``became a passion'' is marked as atomic, as it is a multi-word expression.
%    The dashed line indicates that ``she'' is also a child of the unit ``greatly improved''.
%    UCCA categories, and alignment between source and translation are omitted for brevity.}
%  
%\end{figure*}

@@

@@

\paragraph{Technical Differences}

+ allows for multi-word and even non-contiguous elements

+ allows to indicate the quality of the predicates ("Actions")

\oa{I would consider adding further examples from the Google Doc. For instnace ``your risk of falling'' -- ``Ihr Risiko'' or
  ``Two {\bf review} authors assessed the studies independently and extracted outcome data'' -- ``Zwei Autoren bewertete die Studien unabh\"angig zu {\bf \"uberpr\"ufen}''}

@@

%\begin{figure*}[ht!]
%\begin{tabular}{cc}
%
%
%\subfloat[English-Czech]{
%  \iaafig{cs}
%}
%&
%\subfloat[English-German]{
%  \iaafig{de}
%
%}
%\\
%
%\subfloat[English-Polish]{
%  \iaafig{pl}
%  
%}
%&
%\subfloat[English-Romanian]{
%  \iaafig{ro}
%
%}
%\end{tabular}
%\caption{Confusion matrices for each of the language pairs, plotted as heatmaps.}
%\label{fig:heatmap}
%\end{figure*}
